Schema Change: Challenge Problems

### 2024-05-14 paper.tex
Harsher on Rails

Some tools provide a Domain Specific Language (DSL) to describe schema evolution. Rails Migrations~\cite{RailsMigrations} embeds a DSL in Ruby to describe a range of common schema evolutions. However that falls short of the range of evolutions described by \citet{ambler06}, requiring the addition of custom Ruby or SQL code. \citet{curino08} spawned a stream of research on Database Evolution Languages (DEL) by defining a Schema Modification Operator (SMO) as ``a function that receives as input a relational schema and the underlying database, and produces as output a (modified) version of the input schema and a migrated version of the database''. SMOs can also rewrite queries to accomodate schema changes. \citet{herrmann15} defined a relationally complete DEL and then extended it into a Bidirectional Database Evolution Language (BiDEL)~\cite{herrmann17}. BiDEL appears capable of handling the basic requirements of our \textit{Extract/Absorb Entity} and \textit{Multiplicity Change} challenges, though \textit{Split/Merge Entity} is less clear. It provides schema divergence by supporting multiple schema within one database, but would need some extension to handle data divergence.

>>>>

Some tools provide a Domain Specific Language (DSL) to describe schema evolution. Rails Migrations~\cite{RailsMigrations} embeds a DSL in Ruby to manage schema migration but is comparable to the capabilities of SQL DDL, often requiring the addition of custom Ruby or SQL code. \citet{curino08} spawned a stream of research on Database Evolution Languages (DEL) by defining a Schema Modification Operator (SMO) as ``a function that receives as input a relational schema and the underlying database, and produces as output a (modified) version of the input schema and a migrated version of the database''. SMOs can also rewrite queries to accomodate schema changes. \citet{herrmann15} defined a relationally complete DEL and then extended it into a Bidirectional Database Evolution Language (BiDEL)~\cite{herrmann17}. BiDEL appears capable of handling the basic requirements of our \textit{Extract/Absorb Entity} and \textit{Multiplicity Change} challenges, though \textit{Split/Merge Entity} is less clear. It provides schema divergence by supporting multiple schema within one database, but would need some extension to handle data divergence.
